CITK - an architecture and examples of CUDA enabled ITK filters

### 2011-08-07 article/Article.tex
Small corrections and author modification.

@@

% At minimum, give your name and an email address.  You can include a
% snail-mail address if you like.
\author{Richard Beare{$^1$}, Daniel Micevski, Chris Share\\Luke
  Parkinson, Phil Ward, Wojtek Goscinski{$1$}, Mike Kuiper{$^2$}}
\authoraddress{{$^1$}Richard.Beare@monash.edu, Monash University, Melbourne, Australia\\
{$^2$}mike@vpac.org, Victorian Partnership for Advanced Computing, Melbourne, Australia.}

\begin{document}
\maketitle

@@

The {\em CudaImportImageContainer} manages image data on both the host
and device. When a standard filter requests the image data, such as
through an iterator, the {\em CudaImportImageContainer} checks whether
the most up to date image is on the device or the host. If it is on
the device, it is copied back onto the host. This data is then
supplied to the user. Similarly when a GPU filter requests the image
data, the {\em CudaImportImageContainer} would check where the most up
to date image is, and copy it to the device if required.

@@

The result of this is memory transfers are only performed when
required and are completed transparent to both the developer and the
user. This leaves all the responsibility on the architect, rather than
the developer or the user such as in the other attempts.

@@

\section{Anatomy of a CUDA enabled filter}
CUDA enabled filters can look very like a conventional ITK filter,
with the main difference being a call to a CUDA kernel function from
within the {\em GenerateData} method. CUDA-enabled filters should
never have {\em ThreadedGenerateData} methods as threading is provided
within the CUDA portion.

@@

\section{Performance}
Improved computational performance is the reason for interest in GPU
imaging applications. However there are many stories of how difficult
this is to achieve in practice, and similar difficulties are likely to
be experienced in the imaging domain. Some of the difficulties we
forsee are:
\begin{itemize}
\item A real imaging application is likely to utilize a large number
  of ITK filters. It is likely to be a long time before a significant
  portion of ITK is CUDA (or GPU) enabled. Therefore a developer will
  only experience a nett performance gain if the CUDA enabled filter
  offers sufficient speedup to offset memory transfer costs and a
  small proportion of the application is particularly time consuming.
\item Filters that are easy to port to the GPU tend to be fast on CPU
  anyway, and typically don't represent a large proportion of
  application time. Examples include all voxel-wise operations, such
  as masking and arithmetic.
\item Many potentially time consuming operations, such as filtering
  with large kernels, have been highly optimised for CPU
  implementation. It is important the comparisons are made with these
  optimised CPU implementations.
\end{itemize}

@@

Finally, some positive performance results. These tests were carried
out using a Tesla T10 in a 16 core, 2261.051MHz, Intel L5520  Xeon:
\begin{itemize}
\item Performance improvement of 190 times observed with simple arithmetic, such as adding or subtracting constants from
  images. See {\em simple\_perf\_test.cxx} for details.
\item Image filtering with kernels, for example simple means, offer 32
  times speedup for 3d kernels, radius 10 voxels on images size $500
  \times 500 \times 500$ and 50 times speedup for kernels radius
  20.. This is potentially interesting for applications requiring
  kernels of specific shape, because most accelerated CPU schemes can
  only implement a restricted range of shape kernels. Neither CPU nor
  GPU examples exploits redundancy in this test.
\end{itemize}

>>>>

\newcommand{\IJhandlerIDnumber}{3269}

@@

% At minimum, give your name and an email address.  You can include a
% snail-mail address if you like.
\author{Richard Beare{$^1$}, Daniel Micevski, Chris Share\\Luke
  Parkinson, Phil Ward{$^2$}, Wojtek Goscinski{$1$}, Mike Kuiper{$^3$}}
\authoraddress{{$^1$}Richard.Beare@monash.edu, Monash University, Melbourne, Australia.\\
{$^2$}P.ward@csiro.au, CSIRO Land and Water, Highett, Australia.\\
{$^3$}mike@vpac.org, Victorian Partnership for Advanced Computing, Melbourne, Australia.}

\begin{document}
\IJhandlefooter{\IJhandlerIDnumber}
\maketitle

@@

The {\em CudaImportImageContainer} manages image data on both the host
and device. When a standard filter requests the image data, such as
through an iterator, the {\em CudaImportImageContainer} checks whether
the most up-to-date image is on the device or the host. If it is on
the device, it is copied back onto the host. This data is then
supplied to the user. Similarly when a GPU filter requests the image
data, the {\em CudaImportImageContainer} would check where the most 
up-to-date image is, and copy it to the device if required.

@@

The result of this is memory transfers are only performed when
required and are completely transparent to both the developer and the
user. This leaves all the responsibility on the architect, rather than
the developer or the user such as in the other GPU implementations.

@@

\section{Anatomy of a CUDA enabled filter}
CUDA enabled filters resemble a conventional ITK filter for the most part,
with the main difference being a call to a CUDA kernel function from
within the {\em GenerateData} method. CUDA-enabled filters should
never have {\em ThreadedGenerateData} methods as threading is provided
within the CUDA portion.

@@

\section{Performance}
Improved computational performance is the reason for interest in GPU
imaging applications. However there are many stories of how difficult
this is to achieve in practice, and similar difficulties are likely to
be experienced in the imaging domain. Some of the difficulties we
forsee are:
\begin{itemize}
\item A real imaging application is likely to utilize a large number
  of ITK filters. It is likely to be a long time before a significant
  portion of ITK is CUDA (or GPU) enabled. Therefore a developer will
  only experience a net performance gain if the CUDA enabled filter
  offers sufficient speedup to offset memory transfer costs and a
  small proportion of the application is particularly time consuming.
\item Filters that are easy to port to the GPU tend to be fast on CPU
  anyway, and typically don't represent a large proportion of
  application time. Examples include all voxel-wise operations, such
  as masking and arithmetic.
\item Many potentially time consuming operations, such as filtering
  with large kernels, have been highly optimised for CPU
  implementation. It is important the comparisons are made with these
  optimised CPU implementations.
\end{itemize}

@@

Finally, some positive performance results. These tests were carried
out using a Tesla T10 in a 16 core, 2261.051MHz, Intel L5520  Xeon:
\begin{itemize}
\item Performance improvement of 190 times observed with simple arithmetic, such as adding or subtracting constants from
  images. See {\em simple\_perf\_test.cxx} for details.
\item Image filtering with kernels, for example simple means, offer 32
  times speedup for 3d kernels, radius 10 voxels on images size $500
  \times 500 \times 500$ and 50 times speedup for kernels radius
  20. This is potentially interesting for applications requiring
  kernels of specific shape, because most accelerated CPU schemes can
  only implement a restricted range of shape kernels. Neither CPU nor
  GPU examples exploits redundancy in this test.
\end{itemize}
